An Aliasing-Tolerant Parallel Programming Model

### 2012-11-13 doc/types-pldi2013/alt_example.tex
fixed typos found by Elliott - tried to tweak multicoloring and clobber set paragraphs

Programmers express partitions using {\em colorings},
maps from region elements to integers (``colors'').
Listing~\ref{lst:coloring} shows the function {\tt create\_piece\_coloring}
that constructs a coloring from the piece IDs of a list of nodes.
%The coloring of wires is similar, with each wire
%using the color of its first endpoint.  
For aliased partitions such as the ghost nodes, a {\em multicoloring}
is used, which assigns each region element to one or more colors.  
%(Multicolorings are
%implemented in Core Legion with separate colorings for each color.)
\begin{lstlisting}[label={lst:coloring},caption={Coloring Construction}]
function create_piece_coloring[rl,rn](node_list: NodeList<rl,rn>@rl),
            reads(rl,rn) : coloring(rn) = 
    if isnull(node_list) then
        newcolor(rn) 
    else                                 -- tuple fields accessed by .(field number)
        let list_elem : NodeList<rl,rn> = read(node_list) in
        let part_coloring : coloring(rn) = create_node_coloring(list_elem.2) in
        let node : CircuitNode = read(list_elem.1) in
            color(part_coloring, list_elem.1, node.5)
\end{lstlisting}

>>>>

Programmers express partitions using {\em colorings},
maps from region elements to integers identifying which subregion an element
should be in (``colors'').
Listing~\ref{lst:coloring} shows the function {\tt create\_piece\_coloring}
that constructs a coloring from the piece IDs of a list of nodes.
%The coloring of wires is similar, with each wire
%using the color of its first endpoint.  
For aliased partitions such as the ghost nodes, a {\em multicoloring}
is used, which assigns each region element to one or more colors.  
\footnote{For simplicity, multicolorings are
implemented in Core Legion as separate colorings for each color.}

\begin{lstlisting}[label={lst:coloring},caption={Coloring Construction}]
function create_piece_coloring[rl,rn](node_list: NodeList<rl,rn>@rl),
            reads(rl,rn) : coloring(rn) = 
    if isnull(node_list) then
        newcolor(rn) 
    else                                 -- tuple fields accessed by .(field number)
        let list_elem : NodeList<rl,rn> = read(node_list) in
        let part_coloring : coloring(rn) = create_node_coloring(list_elem.2) in
        let node : CircuitNode = read(list_elem.1) in
            color(part_coloring, list_elem.1, node.5)
\end{lstlisting}

### 2012-11-13 doc/types-pldi2013/langdef.tex
fixed typos found by Elliott - tried to tweak multicoloring and clobber set paragraphs

\infrule[T-Pack]{T_1 = \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1 \\ \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^* \\ \typeenvx e_1 : T_2[r_1/r'_1,\ldots,r_k/r'_k]}{\typeenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] : T_1} &

@@

\infrule[T-Unpack]{T_1 = \exists r'_1, \ldots, r'_n.\ T_2\text{ where }\Omega_1 \andalso \typeenvx e_1 : T_1 \\

@@

  \typeenvx[G=\Gamma',O=\Omega'] e_2 : T_3 \andalso \{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T_1, T_3) = \emptyset}
  {\typeenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 : T_3} &
\infrule[T-Part]{\typeenvx e_1 : \text{coloring}(r_p) \\ \Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{1 \leq i < j \leq k} r_i * r_j \\ \typeenvx[O=\Omega'] e_2 : T \\ \{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T) = \emptyset}

@@

The environment also has a {\em clobber set} $C$.
We will show that parallel Legion programs can be scheduled in
a hierarchical fashion.  We therefore require an
operational semantics that is parallel and hierarchical.  Now
Legion allows limited nondeterminism, specifically because of relaxed coherence on
regions shared between parallel tasks and choices of reduction orderings.
Thus, heap accesses performed by one expression may observe side-effects from 
concurrent evaluations.  The traditional solution to this issue
is to use a top-down, small-step semantics with a global view of all concurrently executing expressions (e.g., \cite{CML}).  
%Rules are specified for which
%expression(s) are valid choices for the next small step of execution.
This approach is incompatible with a hierarchical view in which local
decisions are made with limited information.

@@

\infrule[E-Call]{\opsenvx e_1 \mapsto v_1, E_1 \andalso S_1 = \text{apply}(S, E_1) \andalso \opsenvx e_2 \mapsto v_2, E_2 \andalso \ldots \andalso S_n = \text{apply}(S_{n-1}, E_n) \\

@@

  \opsenvx[M=M',L=L',S=S',C=C'] e_{n+1} \mapsto v_{n+1}, E_{n+1} \andalso E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket) \andalso valid\_interleave(S, C, E'', E', E_{n+1})}

@@

\ruleat{\XA,4.6}{\XB,4.55}{T-Partition}
{\begin{array}{l}
\typeenvx e_1 : \text{coloring}(r_p) \\
\Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{1 \leq i < j \leq k} r_i * r_j \\
\typeenvx[O=\Omega'] e_2 : T \\
\{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T) = \emptyset
\end{array}}
{\typeenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 : T}

\ruleat{0,5.5}{\XB,5.7}{T-Pack}
{\begin{array}{l}
T_1 = \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1 \\
\Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^* \\
\typeenvx e_1 : T_2[r_1/r'_1,\ldots,r_k/r'_k]
\end{array}}
{\typeenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] : T_1}

\ruleat{\XA,6.2}{\XB,6.4}{T-Unpack}
{\begin{array}{l}
T_1 = \exists r'_1, \ldots, r'_n.\ T_2\text{ where }\Omega_1 \\
\typeenvx e_1 : T_1 \\
\Gamma' = \Gamma[T_2[r_1/r'_1,\ldots,r_k/r'_k] / id] \\
\Omega' = \Omega \cup \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \\
\typeenvx[G=\Gamma',O=\Omega'] e_2 : T_3 \\
\{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T_1, T_3) = \emptyset
\end{array}}
{\typeenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 : T_3}

@@

\ruleatx{\XA,8.1}{\XE,8.4}{E-Call}{
\opsenvx e_1 \mapsto v_1, E_1 \andalso S_1 = apply(S, E_1) \andalso \opsenvx e_2 \mapsto v_2, E_2 \andalso \ldots \andalso S_n = apply(S_{n-1}, E_n) \\
valid\_interleave(S, C, E', \oton{E}) \andalso \text{function }id[\otok{k}{r'}](a_1 : T_1, \ldots, a_n : T_n), \Phi', Q' : T_r = e_{n+1} \\
M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \andalso L' = \{ (a_1, v_1), \ldots, (a_n, v_n) \} \andalso S' = apply(S, E') \\
C' = C \cup \{ l : \exists \rho. atomic(\rho) \in M' \llbracket Q' \rrbracket \vee simult(\rho) \in M' \llbracket Q' \rrbracket \} \\
\opsenvx[M=M',L=L',S=S',C=C'] e_{n+1} \mapsto v_{n+1}, E_{n+1} \andalso E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket) \andalso valid\_interleave(S, C, E'', E', E_{n+1})
}{\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E''}

>>>>

\infrule[T-Pack]{T_1 = \exists r'_1, \ldots r'_k.\ T_2\text{ where }\Omega_1 \\ \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^* \\ \typeenvx e_1 : T_2[r_1/r'_1,\ldots,r_k/r'_k]}{\typeenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] : T_1} &

@@

\infrule[T-Unpack]{T_1 = \exists r'_1, \ldots, r'_k.\ T_2\text{ where }\Omega_1 \andalso \typeenvx e_1 : T_1 \\

@@

  \typeenvx[G=\Gamma',O=\Omega'] e_2 : T_3 \andalso \{ \otok{k}{r} \} \cap \textit{regions\_of}(\Gamma, T_1, T_3) = \emptyset}
  {\typeenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}]\ \text{in}\ e_2 : T_3} &
\infrule[T-Part]{\typeenvx e_1 : \text{coloring}(r_p) \\ \Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{1 \leq i < j \leq k} r_i * r_j \\ \typeenvx[O=\Omega'] e_2 : T \\ \{ \otok{k}{r} \} \cap \textit{regions\_of}(\Gamma, T) = \emptyset}

@@

The environment also has a {\em clobber set} $C$, which is the key to handling
limited nondeterminism  while still having an operational
semantics that is parallel and hierarchical.  Nondeterminism arises in a Legion
program because of relaxed coherence on
regions shared between parallel tasks and choices of reduction orderings.
Thus, heap accesses performed by one expression may observe side-effects from 
concurrent evaluations.  The traditional solution to this issue
is to use a top-down, small-step semantics with a global view of all concurrently executing expressions (e.g., \cite{CML}).  
%Rules are specified for which
%expression(s) are valid choices for the next small step of execution.
This approach is incompatible with a hierarchical view in which local
decisions are made with limited information.

@@

\infrule[E-Call]{\opsenvx e_1 \mapsto v_1, E_1 \andalso S_1 = \text{apply}(S, E_1) \andalso \opsenvx[S=S_1] e_2 \mapsto v_2, E_2 \andalso \ldots \andalso S_n = \text{apply}(S_{n-1}, E_n) \\

@@

  \opsenvx[M=M',L=L',S=S',C=C'] e_{n+1} \mapsto v_{n+1}, E_{n+1} \andalso E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket) \andalso valid\_interleave(S, C, E'', E', E'_{n+1})}

@@

\ruleat{\XA,4.6}{\XB,4.55}{T-Partition}
{\begin{array}{l}
\typeenvx e_1 : \text{coloring}(r_p) \\
\Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{1 \leq i < j \leq k} r_i * r_j \\
\typeenvx[O=\Omega'] e_2 : T \\
\{ \otok{k}{r} \} \cap \textit{regions\_of}(\Gamma, T) = \emptyset
\end{array}}
{\typeenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 : T}

\ruleat{0,5.5}{\XB,5.7}{T-Pack}
{\begin{array}{l}
T_1 = \exists r'_1, \ldots r'_k.\ T_2\text{ where }\Omega_1 \\
\Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^* \\
\typeenvx e_1 : T_2[r_1/r'_1,\ldots,r_k/r'_k]
\end{array}}
{\typeenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] : T_1}

\ruleat{\XA,6.2}{\XB,6.4}{T-Unpack}
{\begin{array}{l}
T_1 = \exists r'_1, \ldots, r'_k.\ T_2\text{ where }\Omega_1 \\
\typeenvx e_1 : T_1 \\
\Gamma' = \Gamma[T_2[r_1/r'_1,\ldots,r_k/r'_k] / id] \\
\Omega' = \Omega \cup \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \\
\typeenvx[G=\Gamma',O=\Omega'] e_2 : T_3 \\
\{ \otok{k}{r} \} \cap \textit{regions\_of}(\Gamma, T_1, T_3) = \emptyset
\end{array}}
{\typeenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}]\ \text{in}\ e_2 : T_3}

@@

\ruleatx{\XA,8.1}{\XE,8.4}{E-Call}{
\opsenvx e_1 \mapsto v_1, E_1 \andalso S_1 = apply(S, E_1) \andalso \opsenvx[S=S_1] e_2 \mapsto v_2, E_2 \andalso \ldots \andalso S_n = apply(S_{n-1}, E_n) \\
valid\_interleave(S, C, E', \oton{E}) \andalso \text{function }id[\otok{k}{r'}](a_1 : T_1, \ldots, a_n : T_n), \Phi', Q' : T_r = e_{n+1} \\
M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \andalso L' = \{ (a_1, v_1), \ldots, (a_n, v_n) \} \andalso S' = apply(S, E') \\
C' = C \cup \{ l : \exists \rho. atomic(\rho) \in M' \llbracket Q' \rrbracket \vee simult(\rho) \in M' \llbracket Q' \rrbracket \} \\
\opsenvx[M=M',L=L',S=S',C=C'] e_{n+1} \mapsto v_{n+1}, E_{n+1} \andalso E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket) \andalso valid\_interleave(S, C, E'', E', E'_{n+1})
}{\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E''}
